---
title: 'Assignment: Statistical Inference Course Project 2 Part 1'
author: "John Allen"
date: "Thursday, January 28, 2016"
output:
  pdf_document: default
  html_document:
    keep_md: yes
---
####Load from library
```{r}
library(ggplot2)
```

####Set parameters
```{r}
set.seed(109759)
lambda <- .2
n <- 40
samples <- 1000
```

####Show work and result
```{r}
echo = TRUE
```

####Create a matrix consisting of 1000 simulations of 40 exponentials
```{r}
simulations <- matrix(rexp(n * samples, lambda), samples, n)
simulationsmeans <- rowMeans(simulations)
```

##Show the sample mean and compare it to the theoretical mean of the distribution.

###First lets plot the means
```{r}
hist(main = "Means of the Simulations", simulationsmeans, col = "cyan")
```

####Show the sample mean
```{r}
MEAN <- mean(simulationsmeans)

MEAN
```

####Show the theoretical mean (lambda^-1)
```{r}
THEOMEAN <- lambda^-1

THEOMEAN
```

####There is only a slight difference between the theoretical and sample means of the distribution
```{r}
MEAN - THEOMEAN
```

##Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

####Show the sample variance
```{r}
VARMEAN <- var(simulationsmeans)

VARMEAN
```

####Show the theoretical varicance (lambda * sqrt(n))^-2
```{r}
VARTHEO <- (lambda * sqrt(n))^-2

VARTHEO
```

####Again there is only a small difference between the sample and theoretical
```{r}
VARMEAN - VARTHEO
```

##Show that the distribution is approximately normal

####To accoplish this I will overlay a plot of a normal distribution to a histogram of the sample data
```{r}
plotdata <- data.frame(simulationsmeans);

m <- ggplot(plotdata, aes(x =simulationsmeans))

m <- m + geom_histogram(aes(y=..density..), colour="yellow",
      fill = "red")

m + geom_density(colour="black", size=1);   
```

####This shows that the samples can be approximated with a normal distribution


###

